---
title: "Demonstrating normal distribution models using exponential distribution means"
author: "Michael Fierro"
date: "April 16, 2016"
output: html_document
---

#Synopsis
The

# Report
```{r, echo=FALSE}
set.seed(2291)
lambda_class_1 <- .2
n_sims_class_1 <- 1000
n_obs_class_1 <- 40

initdata <- matrix(rexp(n_sims_class_1 * n_obs_class_1, rate=lambda_class_1),n_sims_class_1,n_obs_class_1)
initMeanData <- rowMeans(initdata)

```
By its definition, an exponential distribution is **not** a normal distribution, and as such, it does not make a gaussian histogram. Instead, it is weighted, going from the highest values at 0 and flattening out to tail along the X axis, as shown with the histogram:
```{r, echo=FALSE}
hist(initdata)
```

However, one would expect a sample of mean values to follow a normal distribution, and the histogram below shows that is exactly what happens:
```{r, echo=FALSE}
hist(initMeanData)
```

### Means
We can now compare the sample and theoretical data:
```{r, echo=FALSE}
paste("Sample mean:",round(mean(initMeanData),3))
paste("Theoretical mean:",round(1/lambda_class_1,3))
```
As expected, the sample mean closely approximates the theoretical mean, being within 0.026. Further samples would bring this even closer, as predicted by the Central Limit Theorem (CLT).

### Variance
The same thing holds true with variances for sample and theoretical data:
```{r, echo=FALSE}
paste("Sample mean:",round(var(initMeanData),3))
paste("Theoretical mean:",round(1/lambda_class_1^2/n_obs_class_1,3))
```
Again, the difference between the variances is negligible and falls perfectly in line with the CLT.

### Distribution

# Appendix
## Code snippets used to generate this report:
```{r, eval=FALSE}
set.seed(2291)
lambda_class_1 <- .2
n_sims_class_1 <- 1000
n_obs_class_1 <- 40

initdata <- matrix(rexp(n_sims_class_1 * n_obs_class_1, rate=lambda_class_1),n_sims_class_1,n_obs_class_1)
initMeanData <- rowMeans(initdata)

```
```{r, eval=FALSE}
hist(initdata)
```
```{r, eval=FALSE}
hist(initMeanData)
```
```{r, eval=FALSE}
paste("Sample mean:",round(mean(initMeanData),3))
paste("Theoretical mean:",round(1/lambda_class_1,3))
```
```{r, eval=FALSE}
paste("Sample mean:",round(var(initMeanData),3))
paste("Theoretical mean:",round(1/lambda_class_1^2/n_obs_class_1,3))
```